---
title: "Creating Quizzes with rquiz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Quizzes with rquiz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(rquiz)
```

## Overview

**rquiz** provides three types of interactive quizzes as HTML widgets:

| Function | Description |
|:---------|:------------|
| `singleQuestion()` | A single question with instant feedback (single- or multiple-choice) |
| `multiQuestions()` | A multi-question quiz with navigation, timer, and results |
| `fillBlanks()` | A fill-in-the-blank cloze text |

All quizzes work in R Markdown, Quarto documents, HTML presentations (e.g. revealJS), and Shiny applications.

## Defining Questions

### Single question (`singleQuestion`)

Questions are defined as a named list with 3 (or 4) elements:

```{r, eval = FALSE}
myQuestionSC <- list(
  question = "Which city hosted the 2024 Summer Olympics?", # the question text
  options = c("Berlin", "London", "Paris", "Madrid"),  # answer options
  answer = 3,  # index of the correct answer (Paris)
  tip = "Think about the Eiffel Tower."  # optional tip text
)
```

For **single-choice**, `answer` is a single integer. For **multiple-choice**, `answer` is a vector of integers. The quiz mode is auto-detected from the length of `answer`:

```{r, eval = FALSE}
myQuestionMC <- list(
  question = "Which of the following are prime numbers?",
  options = c("2", "4", "7", "9", "11", "15"),
  answer = c(1, 3, 5),  # indices of correct options (2, 7, 11)
  tip = "Prime numbers are only divisible by 1 and themselves."
)
```

You can validate your question with `checkSingleQuestion()` before passing it to the quiz function. This catches common mistakes like missing fields, out-of-range indices, or misspelled element names:

```{r, eval = FALSE}
checkSingleQuestion(myQuestionSC)  # returns invisibly if valid

# Common mistake: unnamed or partially named list
bad <- list("What is 2+2?", c("3", "4"), answer = 2)
checkSingleQuestion(bad)
#> Error: The question list must have named elements: 'question', 'options',
#>   and 'answer' (plus optionally 'tip'). Missing: question, options.
#>   Example: list(question = "What is 2+2?", options = c("3", "4", "5"), 
#>   answer = 2)
#>   See ?checkSingleQuestion or ?singleQuestion for examples.
```

### Multiple questions (`multiQuestions`)

Pass a list of question lists. Each question follows the same structure. Sub-lists can be named (e.g. `q1`, `q2`) for readability, but this is not required.

**Single-choice only** - all questions have a single correct answer:

```{r, eval = FALSE}
sportQuiz <- list(
  q1 = list(
    question = "Which city hosted the 2024 Summer Olympics?",
    options = c("Athens", "London", "Paris", "Tokyo"),
    answer = 3
  ),
  q2 = list(
    question = "Which sport was added to the Olympics at the 2024 Paris Games?",
    options = c("Skateboarding", "Breakdance", "Surfing", "Sport Climbing"),
    answer = 2
  ),
  q3 = list(
    question = "At the 1900 Olympics in Paris, one sport was featured that has never returned to the Games since. Which one?",
    options = c("Live pigeon shooting", "Underwater archery", "Synchronized sneezing", 
      "Backwards horse vaulting"),
    answer = 1
  )
)
```

**Mixed with tips** - some questions include a custom tip text:

```{r, eval = FALSE}
sportQuizWithTips <- list(
  q1 = list(
    question = "Which city hosted the 2024 Summer Olympics?",
    options = c("Athens", "London", "Paris", "Tokyo"),
    answer = 3,
    tip = "They speak French there."
  ),
  q2 = list(
    question = "Which sport was added to the Olympics at the 2024 Paris Games?",
    options = c("Skateboarding", "Breakdance", "Surfing", "Sport Climbing"),
    answer = 2
    # no tip for this question - the tip button won't appear here
  )
)
```

**Multiple-choice** - if **any** question has multiple correct answers (`length(answer) > 1`), all questions automatically use checkboxes:

```{r, eval = FALSE}
scienceQuiz <- list(
  q1 = list(
    question = "Which of these are noble gases?",
    options = c("Helium", "Oxygen", "Neon", "Iron"),
    answer = c(1, 3)
  ),
  q2 = list(
    question = "Which planet is closest to the Sun?",
    options = c("Venus", "Mercury", "Mars"),
    answer = 2  # single correct answer, but rendered as checkbox
  )
)
```

Validate the entire question list with `checkMultiQuestions()`:

```{r, eval = FALSE}
checkMultiQuestions(sportQuiz)  # returns invisibly if valid

# Common mistake: flat list instead of list of lists
bad <- list(question = "Q1?", options = c("A", "B"), answer = 1,
            question = "Q2?", options = c("C", "D"), answer = 2)
checkMultiQuestions(bad)
#> Error: `x` looks like a single question, not a list of questions.
#>   Wrap it in an outer list: list(q1 = list(question = ..., options = ..., 
#>   answer = ...))
#>   See ?checkMultiQuestions or ?multiQuestions for examples.
```

### Fill-in-the-blank (`fillBlanks`)

Define a cloze text with `$$!answer!$$` markers for blank fields. Optionally add distractor options:

```{r, eval = FALSE}
rQuiz <- list(
  cloze = "R is a $$!programming!$$ language for $$!statistical computing!$$.",
  addOptions = c("natural", "colloquial", "movies")
)
```

Normal text wraps automatically across lines. For code formatting, use `<pre>` tags — but note that inside `<pre>`, you need `<br>` tags for explicit line breaks:

```{r, eval = FALSE}
codeQuiz <- list(
  cloze = "<pre>x $$!<-!$$ c(1, 2, 3)
$$!mean!$$(x)</pre>",
  addOptions = c("=", "->", "median", "sum")
)
```

Validate with `checkFillBlanks()`:

```{r, eval = FALSE}
checkFillBlanks(rQuiz)  # returns invisibly if valid

# Common mistake: misspelled element name
bad <- list(clozed = "R is a $$!programming!$$ language.")
checkFillBlanks(bad)
#> Error: Unknown element(s) in `x`: 'clozed'.
#>   Allowed elements are: 'cloze' (required) and 'addOptions' (optional).
#>   See ?checkFillBlanks or ?fillBlanks for examples.
```

## Creating a Quiz

Once your data is defined, pass it to the quiz function:

```{r}
myQuestionSC <- list(
  question = "Which city hosted the 2024 Summer Olympics?", # the question text
  options = c("Berlin", "London", "Paris", "Madrid"),  # answer options
  answer = 3,  # index of the correct answer (Paris)
  tip = "Think about the Eiffel Tower."  # optional tip text
)
singleQuestion(
  x = myQuestionSC,
  title = "Geography Quiz",
  showTipButton = TRUE
)
```

## Key Parameters

All three quiz functions share a common set of parameters:

### Content & behavior

| Parameter | Description | Default |
|:----------|:------------|:--------|
| `title` | Quiz title text | `NULL` / `"Quiz"` |
| `language` | Language for UI text (`"en"`, `"de"`, `"fr"`, `"es"`) | `"en"` |
| `shuffle` | Randomize option/question order | `FALSE` |
| `showTipButton` | Show a hint button | `FALSE` |
| `showSolutionButton` | Show a reveal-solution button | `TRUE` |

**Tips:** Set `showTipButton = TRUE` to display a tip button. The behavior depends on the quiz type and mode:

- **Single-choice (`singleQuestion` and `multiQuestions`):** Tips are only shown if you provide a custom tip text as the 4th element of the question list (`tip = "..."`). Questions without a tip simply don't show the button.
- **Multiple-choice:** If no custom tip is provided, an automatic tip is generated showing the number of correct answers (e.g. "Number of correct answers: 3"). You can still override this with your own tip text.
- **Fill-in-the-blank (`fillBlanks`):** The tip shows all available answer options (correct answers + distractors from `addOptions`).

**Submit & retry (MC mode in `singleQuestion`):** After clicking Submit, a feedback message shows how many correct and wrong selections were made. The Submit button is disabled until the user changes their selection (adds or removes an option), then it can be clicked again. This allows iterative learning — adjust your answer and retry without reloading the page.

**Shuffling:** Set `shuffle = TRUE` to randomize order:

- **`singleQuestion`:** The answer options are displayed in a random order each time the widget is rendered (e.g. when the page is loaded or refreshed).
- **`multiQuestions`:** The question order is randomized on each render, and also each time the user clicks "Try again" on the results page.

### Layout

| Parameter | Description | Default |
|:----------|:------------|:--------|
| `width` | Widget width (CSS value) | `"100%"` |
| `height` | Widget height (CSS value) | `"500px"` |
| `scroll` | Fixed height with scrollbar (useful for slides) | `FALSE` |
| `center` | Center the widget horizontally | `TRUE` |

### Design

All visual aspects are customizable --- colors, fonts, sizes. Common parameters include:

| Parameter | Description | Default |
|:----------|:------------|:--------|
| `fontFamily` | Font stack | `"'Helvetica Neue', ..."` |
| `fontSize` | Base font size in px | `16` |
| `titleCol` / `titleBg` | Title text & background color | `"#FFF"` / `"#5F5F5F"` |
| `questionCol` / `questionBg` | Question (or description in fillBlanks) text & background color. Also serves as feedback area (green/red). | varies |

See the function documentation (`?singleQuestion`, `?multiQuestions`, `?fillBlanks`) for the full list of design parameters.

## Reusable Themes

If you use multiple quizzes in the same document and want a consistent look, create a reusable theme with `rquizTheme()` instead of repeating color settings:

```{r, eval = FALSE}
# Define a dark theme once
dark <- rquizTheme(
  fontFamily = "Georgia, serif",
  fontSize = 18,
  titleCol = "#E0E0E0", titleBg = "#1A1A2E",
  questionCol = "#FFFFFF", questionBg = "#16213E",
  mainCol = "#E0E0E0", mainBg = "#1A1B2E",
  optionBg = "#252540",
  navButtonCol = "#FFFFFF", navButtonBg = "#E94560",
  tipButtonCol = "#E0E0E0", tipButtonBg = "#2C2C3E",
  solutionButtonCol = "#E0E0E0", solutionButtonBg = "#2C2C3E"
)

# Apply to any quiz - all share the same design:
singleQuestion(x = myQuestionSC, theme = dark, title = "Quiz 1")
multiQuestions(x = scienceQuiz, theme = dark, title = "Quiz 2")
fillBlanks(x = rQuiz, theme = dark, title = "Quiz 3")
```

The theme provides defaults that are overridden by any explicitly passed arguments:

```{r, eval = FALSE}
# Use the dark theme but override the title background:
singleQuestion(x = myQuestionSC, theme = dark, title = "Quiz 1", 
  titleBg = "#FF0000")
```

See `?rquizTheme` for all available theme parameters, including quiz-specific ones like `optionLabelBg` (singleQuestion), `navButtonBg` (multiQuestions), or `descriptFontSize` (fillBlanks).

## Formatting Text with HTML

Since quiz content is rendered as HTML, you can use HTML tags to format text within questions, options, tips, and cloze text. Standard Markdown formatting (`*italic*`, `**bold**`) does **not** work inside quiz strings — use HTML tags instead.

### Italic and bold text

Use `<em>` for italics (e.g. species names) and `<strong>` for bold:

```{r, eval = FALSE}
list(
  question = "What is the scientific name of the blue whale?",
  options = c(
    "<em>Balaenoptera musculus</em>",
    "<em>Megaptera novaeangliae</em>",
    "<em>Physeter macrocephalus</em>"
  ),
  answer = 1
)
```

### Inline code

Use `<code>` for inline code formatting:

```{r, eval = FALSE}
list(
  question = "What does the pipe operator <code>|></code> do in R?",
  options = c(
    "It performs logical OR operations",
    "It passes the left side as the first argument to the right side",
    "It creates a new variable"
  ),
  answer = 2
)
```

### Colored text

Use `<span>` with inline CSS for colored or styled text:

```{r, eval = FALSE}
list(
  question = "Which <span style='color: #E74C3C;'>color</span> does litmus paper turn in an acidic solution?",
  options = c("Blue", "Red", "Green", "Yellow"),
  answer = 2
)
```

### Code formatting in fill-in-the-blank quizzes

Use `<pre>` tags for monospace code blocks in `fillBlanks()`. Inside `<pre>`, line breaks require the `<br>` tag (normal line breaks are ignored):

```{r, eval = FALSE}
list(
  cloze = "<pre>library($$!ggplot2!$$)<br>
ggplot(data = iris) +<br>
  $$!geom_point!$$(aes(x = Sepal.Length, y = Sepal.Width))</pre>",
  addOptions = c("geom_line", "dplyr")
)
```


## Multilingual Support

Set `language` to change all UI text (buttons, messages) automatically:

```{r, eval = FALSE}
# German
singleQuestion(x = myQuestionSC, language = "de")

# French
multiQuestions(x = sportQuiz, language = "fr")

# Spanish
fillBlanks(x = rQuiz, language = "es")
```

Currently supported: English (`"en"`), German (`"de"`), French (`"fr"`), Spanish (`"es"`). To request additional languages, please open an issue on [GitHub](https://github.com/saskiaotto/rquiz/issues).

## Using in Quarto / R Markdown

Simply include quiz code in a code chunk. The widget renders as part of the HTML output:

````markdown
```{r}`r ''`
library(rquiz)
singleQuestion(
  x = list(
    question = "What is 2+2?",
    options = c("3", "4", "5"),
    answer = 2
  ),
  title = "Math"
)
```
````

For **HTML slide presentations** (e.g. revealJS, xaringan), use `scroll = TRUE` to add a scrollbar when the quiz exceeds the available slide space:

```{r, eval = FALSE}
singleQuestion(x = myQuestionSC, scroll = TRUE, height = "400px")
```

## Using in Shiny

Use the provided output/render functions:

```{r, eval = FALSE}
library(shiny)
library(rquiz)

ui <- fluidPage(
  h2("My Quiz App"),
  singleQuestionOutput("quiz")
)

server <- function(input, output) {
  output$quiz <- renderSingleQuestion({
    singleQuestion(
      x = list(
        question = "What is 2+2?",
        options = c("3", "4", "5"),
        answer = 2
      )
    )
  })
}

shinyApp(ui, server)
```

Each quiz type has its own pair: `singleQuestionOutput()` / `renderSingleQuestion()`, `multiQuestionsOutput()` / `renderMultiQuestions()`, `fillBlanksOutput()` / `renderFillBlanks()`.

## Next Steps

- Browse the [Gallery](gallery.html) for interactive examples with different configurations
- Check the [Function Reference](../reference/index.html) for all available parameters
